A User-Extensible and Adaptable Parser Architecture

### 2008-08-20 doc/sgai-2008.tex
Corrections from the second half of the paper.

\title*{A user-extensible and adaptable parser architecture}
\author{John Tobin and Carl Vogel}
\institute{John Tobin \at{} School of Computer Science and Statistics,
Trinity College, Dublin 2, Ireland. \newline{} \email{tobinjt@cs.tcd.ie}
\and{} Carl Vogel \at{} School of Computer Science and Statistics,
Trinity College, Dublin 2, Ireland. \newline{} \email{vogel@cs.tcd.ie}}
\maketitle

@@

\begin{table}[hbtp]
    \caption{Similarities with ATN}\label{Similarities with ATN}
    % The \smallskip{} below stop the text hitting the lines.
    \begin{tabular}[]{lll}
        \tabletopline{}
        ATN           & Architecture    & Similarity                                \\
        \tablemiddleline{}
        Networks      & Framework       & Determines the sequence of transitions or \\
                      &                 & actions which constitutes a valid input.  \\
        Transitions   & Actions         & Save data and impose conditions the       \\
                      &                 & input must meet to be considered valid.   \\
        Abbreviations & Rules           & Responsible for classifying input.        \\
        \tablebottomline{}
    \end{tabular}
\end{table}

@@

Parsing efficiency is an obvious concern when the Postfix log parser
routinely needs to parse very large log files.  The mail server which
generated the log files used in testing the Postfix log parser accepts
approximately 10,000 mails for 700 users a day.  Median log file size is
50MB, containing 285,000 log lines --- large scale mail servers would have
much larger log files.  When generating the timing data used in this
section, \numberOFlogFILES{} log files (totaling 10.08 GB,
\numberOFlogLINEShuman{} log lines) were each parsed 10 times and the
parsing times averaged.  Saving results to the database was disabled for
the test runs, as the tests are aimed at measuring the speed of the Postfix
log parser rather than the speed of the database and the disks the database
is stored on.  The computer used for test runs was a Dell Optiplex 745
described in \tableref{Computer used to generate statistics}.  Parsing all
\numberOFlogFILES{} log files in one run took
\input{build/include-timing-run-duration}.  The computer was dedicated to
the task of gathering statistics from test runs, and was not used for any
other purpose while test runs were ongoing; any services not necessary for
running the tests were disabled.

% XXX MAYBE CHANGE POSITIONING
\begin{table}[htbp]
    \caption{Computer used to generate statistics}
    \empty{}\label{Computer used to generate statistics}
    \begin{tabular}[]{ll}
        \tabletopline{}
        Component  & Component used                                 \\
        \tablemiddleline{}
        CPU        & 1 dual core 2.40GHz Intel\textregistered{}
                     Core\texttrademark{}2 CPU,                      
                     with 32KB L1 and 4MB L2 cache.                 \\
        RAM        & 2GB 667 MHz DDR RAM\@.                         \\
        Hard disk  & 1 Seagate Barracuda 7200 RPM 250GB SATA disk.  \\
        \tablebottomline{}
    \end{tabular}
\end{table}

@@

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{Parsing time, file size, and number of lines} shows the parsing
time in seconds, file size in MB, and number of log lines in tens of
thousands, for each of the \numberOFlogFILES{} log files.  All three lines
run roughly in parallel, giving the impression that the algorithm scales
linearly with input size.  This impression is borne out by
\graphref{parsing time vs file size vs number lines factor}: the ratios are
quite tightly banded across the graph, showing that the algorithm scales
linearly.  The ratio increases (i.e.\ improves) for log files 22 and 62--68
despite their large size (due to a mail forwarding loop
resulting in a greatly increased number of mails delivered and log lines
generated).

@@

At the time of writing there are \numberOFrules{} different rules, with the
top 10\% matching the vast majority of the log lines, and the remaining log
lines split across the other 90\% of the rules in a Power Law distribution
(as shown in \graphref{rule hits graph}).  Assuming that the distribution
of log lines is reasonably steady over time, parser efficiency should
benefit from trying more frequently matching rules before those which match
less frequently.  To test this hypothesis three full test runs were
performed with different rule orderings:

@@

    \item [optimal]  Hypothetically the most optimal order: rules which
        match most often will be tried first.

    \item [shuffle] Random ordering, intended to represent an unsorted rule
        set.  The rules will be shuffled once before use and will retain
        that ordering for the entirety of the log file.  Note that the
        ordering will change every time the parser is executed, so 10
        different orderings will be generated for each log file in the test
        run.

@@

\Graphref{Parsing time relative to shuffled} shows the parsing times of
optimal and reversed orderings as a percentage of the parsing time of
shuffled ordering.  Overall this optimisation provides a modest but
worthwhile reduction in parsing time of approximately 10\%.
\showgraph{build/graph-hits}{Hits per rule}{rule hits graph}
\showgraph{build/graph-optimal-and-reverse-vs-shuffle}{Parsing time
relative to shuffled ordering}{Parsing time relative to shuffled}

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that there will be a large number of rules.  There are
\numberOFrules{} rules in the full Postfix log parser ruleset, whereas the
minimum number of rules required to parse the \numberOFlogFILES{} log files
is \numberOFrulesMINIMUM{}, \numberOFrulesMINIMUMpercentage{} of the full
ruleset.  A second set of statistics was generated using the minimum
ruleset and the parsing times compared to those generated using the full
ruleset: the percentage parsing time increase when using the full ruleset
instead of the minimal ruleset for optimal, shuffled and reversed orderings
is shown in \graphref{Percentage parsing time increase of maximum ruleset
over minimum ruleset}.  It is clear from \graphref{Percentage parsing time
increase of maximum ruleset over minimum ruleset} that the increased number
of rules has a noticeable performance impact with reverse ordering, and a
lesser impact with shuffled ordering.  The optimal ordering shows a mean
increase of \input{build/include-full-ruleset-vs-minimum-ruleset} in
parsing time for a \numberOFrulesMAXIMUMpercentage{} increase in the number
of rules.  These results show that the architecture scales very well as the
number of rules increases, and that sorting the rules is an important
optimisation enabling this.
\showgraph{build/graph-full-ruleset-vs-minimum-ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}

@@

The Postfix log parser's coverage of log files is separated into two parts:
log line coverage and mail coverage (the former is a requirement for the
latter to be achieved).  Improving the former is less intrusive, as it just
requires new rules to be written; improving the latter is more intrusive as
it requires changes to actions, and it can also be much harder to notice a
deficiency.

Warnings are issued for any log lines which are not parsed; no warnings are
issued for unparsed log lines while parsing the \numberOFlogFILES{} test
log files, so it can be safely concluded that there are zero false
negatives.  False positives are harder to quantify: manually verifying that
the correct regex parsed each of the \numberOFlogLINES{} log lines is
infeasible, making it impossible to quantify the false positive rate.  A
random sample of 6039 log lines (0.00994\% of the total) was parsed and the
results manually verified by inspection to ensure that the correct regex
parsed each log line.  The sample contained zero false positives, and this
check has been automated to ensure continued accuracy.

Coverage of mails is much more difficult to determine accurately than
coverage of log lines.  The implementation can dump its state tables in a
human readable form; examining these tables with reference to the log files
is the best way to detect mails which were not handled properly.  The
implementation issues warnings when it detects any errors, some of which
may alert the user to a problem.  There should be few or no warnings when
parsing, and when finished parsing the state table should only contain
entries for mails which had yet to be delivered when the log files ended,
or were accepted before the log files began.  A second sample of 6000 log
lines was parsed with all debugging options enabled, resulting in 167,448
lines of output.  All 167,448 lines were examined in conjunction with the
input log file and a dump of the resulting database, verifying that for
each of the input lines the implementation used the correct rule and
executed the correct action, which in turn produced the correct result and
inserted the correct data in the database.  The log segment produced 4
warnings (about mails which had started before the log file segment), 10
mails remaining in the state tables, and 1625 connections correctly entered
in the database.  No error messages were produced, therefore there were no
false negatives.  Given the evidence detailed above, the authors are
confident that the false positive rate when reconstructing a mail is zero.

@@

The architecture's greatest strength is the ease with which it can be
adapted to deal with new requirements and inputs.  Parsing a variation of
an existing input is a trivial task: simply modify an existing rule or add
a new one with an appropriate regex and the task is complete; knowledge of
the framework is not required when writing rules.  Parsing a new category
of input is achieved by writing a new action and appropriate rules; quite
often the action will not need to interact directly with existing actions,
but when interaction is required the framework provides the necessary
facilities.  The decoupling of rules from actions allows different sets of
rules to be used with the same actions: e.g.\ a parser might have actions
to process versions one and two of a file format; the parser will then
parse either version one, version two, or both versions depending on the
ruleset selected.  The architecture makes it possible to apply commonly
used programming techniques such as object orientation, inheritance,
composition, delegation and roles when designing and implementing a parser,
simplifying the process of working within a team with shared
responsibility, or when developing and testing additional functionality.
This architecture is ideally suited to parsing inputs where the input is
not fully understood or does not follow a fixed grammar: the architecture
warns about unparsed inputs and other errors, but continues on parsing as
best it can, allowing the developer of a new parser to decide which
deficiencies are most important and require attention first, rather than
being forced to fix the first error which arises.

The dataset gathered by the Postfix log parser provides the foundation for
the future of this project: applying machine-learning algorithms to the
data to analyse and optimise the set of rules in use, followed by
identifying patterns in the data which could be used to write new filters
to recognise and reject spam rather than accepting it.  The parser provides
the data in a normalised form which is far easier to use as input to new or
existing implementations of machine-learning algorithms than trying to
adapt each algorithm to extract data directly from the log files.  The
current focus is on clustering and decision trees to optimise the order in
which rules are applied; future efforts will involve using data gathered by
the parser to train and test new filters.  This task is very similar to
optimising a black box application based on its inputs and outputs, and
this approach could be applied to optimising the behaviour of any system
given sufficient logging to analyse.  An alternative approach to black box
optimisation which uses application profiling in conjunction with the
application's error messages to improve the error messages shown to users
is described in~\cite{black-box-error-reporting}; profiling data may be
useful in supplementing systems which fail to provide adequate log
messages.

The Postfix log file parser based on this architecture provides a basis for
systems administrators to monitor the effectiveness of their anti-spam
measures and adapt their defences to combat the new techniques used by
those sending spam.  The parser is a fully usable application, built to
address a genuine need, rather than a proof of concept whose sole purpose
is to illustrate a new idea; it deals with the oddities and difficulties
which occur in the real world, rather than a clean, idealised scenario
developed to showcase the best features of a new approach.

>>>>

\title*{A User-Extensible and Adaptable Parser Architecture}
\author{John Tobin and Carl Vogel}
\institute{John Tobin \at{} School of Computer Science and Statistics,
Trinity College, Dublin 2, Ireland. \newline{} \email{tobinjt@cs.tcd.ie}
\and{} Carl Vogel \at{} School of Computer Science and Statistics,
Trinity College, Dublin 2, Ireland. \newline{} \email{vogel@cs.tcd.ie}}
\maketitle

@@

\begin{table}[hbtp]
    \caption{Similarities with ATN}\label{Similarities with ATN}
    % The \smallskip{} below stop the text hitting the lines.
    \begin{tabular}[]{lll}
        \tabletopline{}%
        ATN           & Architecture    & Similarity                                \\
        \tablemiddleline{}%
        Networks      & Framework       & Determines the sequence of transitions or \\
                      &                 & actions which constitutes a valid input.  \\
        Transitions   & Actions         & Save data and impose conditions the       \\
                      &                 & input must meet to be considered valid.   \\
        Abbreviations & Rules           & Responsible for classifying input.        \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

Parsing efficiency is an obvious concern when the Postfix log parser
routinely needs to parse very large log files.  The mail server which
generated the log files used in testing the Postfix log parser accepts
approximately 10,000 mails for 700 users a day.  Median log file size is
50MB, containing 285,000 log lines --- large scale mail servers would have
much larger log files.  When generating the timing data used in this
section, \numberOFlogFILES{} log files (totaling 10.08 GB,
\numberOFlogLINEShuman{} log lines) were each parsed 10 times and the
parsing times averaged.  Saving results to the database was disabled for
the test runs, as the tests are aimed at measuring the speed of the Postfix
log parser rather than the speed of the database and the disks the database
is stored on.  The computer used for test runs was a Dell Optiplex 745
described in \tableref{Computer used to generate statistics}; it was
dedicated to the task of gathering statistics from test runs, and was not
used for any other purpose while test runs were ongoing --- any services
not necessary for running the tests were disabled.  Parsing all
\numberOFlogFILES{} log files in one run took
\input{build/include-timing-run-duration}.

% XXX MAYBE CHANGE POSITIONING
\begin{table}[htbp]
    \caption{Computer used to generate statistics}
    \empty{}\label{Computer used to generate statistics}
    \begin{tabular}[]{ll}
        \tabletopline{}%
        Component  & Component used                                 \\
        \tablemiddleline{}%
        CPU        & 1 dual core 2.40GHz Intel\textregistered{}
                     Core\texttrademark{}2 CPU,                      
                     with 32KB L1 and 4MB L2 cache.                 \\
        RAM        & 2GB 667 MHz DDR RAM\@.                         \\
        Hard disk  & 1 Seagate Barracuda 7200 RPM 250GB SATA disk.  \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

An important property of a parser is how parsing time scales relative to
input size: linearly, polynomially, or exponentially?  \Graphref{Parsing
time, file size, and number of lines} shows the parsing time in seconds,
file size in MB, and number of log lines in tens of thousands, for each of
the \numberOFlogFILES{} log files.  All three lines run roughly in
parallel, giving the impression that the algorithm scales linearly with
input size.  This impression is borne out by \graphref{parsing time vs file
size vs number lines factor}: the ratios are tightly banded across the
graph, showing that the algorithm scales linearly.  The ratios increase
(i.e.\ improve) for log files 22 and 62--68 despite their large size due to
a mail forwarding loop resulting in a greatly increased number of mails
delivered and log lines generated.

@@

At the time of writing the Postfix log parser has \numberOFrules{} different rules, with the
top 10\% matching the vast majority of log lines, and the remaining log
lines split across the other 90\% of the rules in a Power Law distribution
(as shown in \graphref{rule hits graph}).  Assuming that the distribution
of log lines is reasonably consistent over time, parser efficiency should
benefit from trying more frequently matching rules before those which match
less frequently.  To test this hypothesis three full test runs were
performed with different rule orderings:

@@

    \item [optimal]  Hypothetically the best order: rules which match most
        often will be tried first.

    \item [shuffle] Random ordering, intended to represent an unsorted rule
        set.  Note that the ordering will change every time the parser is
        executed, so 10 different orderings will be generated for each log
        file in the test run.

@@

\Graphref{Parsing time relative to shuffled} shows the parsing times of
optimal and reversed orderings as a percentage of the parsing time of
shuffled ordering.  Overall this optimisation provides a modest but
worthwhile reduction in parsing time of approximately 10\%.  Optimal rule
ordering is also beneficial in other ways, as shown in
\sectionref{Architecture Scalability: Number of Rules}.
\showgraph{build/graph-hits}{Hits per rule}{rule hits graph}
\showgraph{build/graph-optimal-and-reverse-vs-shuffle}{Parsing time
relative to shuffled ordering}{Parsing time relative to shuffled}

@@

\label{Architecture Scalability: Number of Rules}

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that typical parsers will have a large number of rules.  There are
\numberOFrules{} rules in the full Postfix log parser ruleset, whereas the
minimum number of rules required to parse the \numberOFlogFILES{} log files
is \numberOFrulesMINIMUM{}, \numberOFrulesMINIMUMpercentage{} of the full
ruleset.  A second set of statistics was generated using the minimum
ruleset and the parsing times compared to those generated using the full
ruleset: the percentage parsing time increase when using the full ruleset
instead of the minimal ruleset for optimal, shuffled and reversed orderings
is shown in \graphref{Percentage parsing time increase of maximum ruleset
over minimum ruleset}.  It is clear that the increased number of rules has
a noticeable performance impact with reverse ordering, and a lesser impact
with shuffled ordering.  The optimal ordering shows a mean increase of
\input{build/include-full-ruleset-vs-minimum-ruleset} in parsing time for a
\numberOFrulesMAXIMUMpercentage{} increase in the number of rules.  These
results show that the architecture scales very well as the number of rules
increases, and that sorting the rules is an important optimisation enabling
this.  \showgraph{build/graph-full-ruleset-vs-minimum-ruleset}{Percentage
parsing time increase of maximum ruleset over minimum ruleset}{Percentage
parsing time increase of maximum ruleset over minimum ruleset}

@@

The Postfix log parser requires two different types of coverage to be
measured: log lines parsed and mail delivery attempts correctly understood
(the former is a requirement for the latter to be achieved).  Improving the
former is less intrusive, as it just requires new rules to be written;
improving the latter is more intrusive as it requires changes to actions,
and it can also be much harder to notice a deficiency.

Measuring Postfix log parser coverage begins with log line coverage.
Warnings are issued for any log lines which are not parsed; no such
warnings are issued while parsing the \numberOFlogFILES{} test log files,
so it can be safely concluded that there are zero false negatives.  False
positives are harder to quantify: manually verifying that the correct rule
parsed each of the \numberOFlogLINES{} log lines is infeasible, making it
impossible to quantify the false positive rate.  A random sample of 6039
log lines (0.00994\% of the total) was parsed and the results manually
verified by inspection to ensure that the correct rule parsed each log
line.  The sample contained zero false positives, and this check has been
automated to ensure continued accuracy.

The proportion of mail delivery attempts correctly understood is much more
difficult to determine accurately than coverage of log lines.  The
implementation can dump its state tables in a human readable form;
examining these tables with reference to the log files is the best way to
detect mails which were not handled properly.  The implementation issues
warnings when it detects any errors, which should alert the user to a
problem.  There should be few or no warnings when parsing, and when
finished parsing the state table should only contain entries for mails
which had yet to be delivered when the log files ended, or were accepted
before the log files began.  A second sample of 6000 log lines was parsed
with all debugging options enabled, resulting in 167,448 lines of output.
All 167,448 lines were examined in conjunction with the input log file and
a dump of the resulting database, verifying that for each of the input
lines the implementation used the correct rule and executed the correct
action, which in turn produced the correct result and inserted the correct
data in the database.  The log segment produced 4 warnings (about mails
which had started before the log file segment), 10 mails remaining in the
state tables, and 1625 connections correctly entered in the database.  No
error messages were produced, therefore there were no false negatives.
Given the evidence detailed above, the authors are confident that the false
positive rate when reconstructing a mail is zero.

Experience implementing the Postfix log parser shows that full input
coverage is relatively easy to achieve with this architecture, and that
with sufficient time and effort full understanding of the input is
possible.  Postfix log files would require substantial time and effort to
correctly parse regardless of the architecture used; this architecture
enables an iterative approach to be used (similar to Stepwise
Refinement~\cite{stepwise-refinement}), as is practiced in many other
software engineering disciplines.

@@

The architecture's greatest strength is the ease with which it can be
adapted to deal with new requirements and inputs.  Parsing a variation of
an existing input is a trivial task: simply modify an existing rule or add
a new one with an appropriate regex and the task is complete; knowledge of
the framework is not required.  Parsing a new category
of input is achieved by writing a new action and appropriate rules; quite
often the action will not need to interact directly with existing actions,
but when interaction is required the framework provides the necessary
facilities.  The decoupling of rules from actions allows different sets of
rules to be used with the same actions, e.g.\ a parser might have actions
to process versions one and two of a file format; the parser will then
parse either version one, version two, or both versions, depending on the
ruleset selected.  The architecture makes it possible to apply commonly
used programming techniques such as object orientation, inheritance,
composition, delegation, and roles when designing and implementing a parser,
simplifying the process of working within a team with shared
responsibility for a parser, or when developing and testing additional functionality.
This architecture is ideally suited to parsing inputs where the input is
not fully understood or does not follow a fixed grammar: the architecture
warns about unparsed inputs and other errors, but continues on parsing as
best it can, allowing the developer of a new parser to decide which
deficiencies are most important and require attention first, rather than
being forced to fix the first error which arises.

The dataset gathered by the Postfix log parser provides the foundation for
the future of this project: applying machine-learning algorithms to the
data to analyse and optimise the set of rules in use, followed by
identifying patterns in the data which could be used to write new filters
to recognise and reject spam rather than accepting it.  The parser provides
the data in a normalised form which is far easier to use as input to new or
existing implementations of machine-learning algorithms than trying to
adapt each algorithm to extract data directly from the log files.  The
current focus is on clustering and decision trees to optimise the order in
which rules are applied; future efforts will involve using data gathered by
the parser to train and test new filters.  This task is very similar to
optimising a black box application based on its inputs and outputs, and
this approach could be applied to optimising the behaviour of any system
given sufficient log messages to analyse.  An alternative approach to black box
optimisation which uses application profiling in conjunction with the
application's error messages to improve the error messages shown to users
is described in~\cite{black-box-error-reporting}; profiling data may be
useful in supplementing systems which fail to provide adequate log
messages.

The Postfix log file parser based on this architecture provides a basis for
systems administrators to monitor the effectiveness of their anti-spam
measures and adapt their defences to combat new techniques used by those
sending spam.  This parser is a fully usable application, built to address
a genuine need, rather than a proof of concept whose sole purpose is to
illustrate a new idea; it deals with the oddities and difficulties which
occur in the real world, rather than a clean, idealised scenario developed
to showcase the best features of a new approach.
